---
title: "a3_task3_gabrielle_barsotti"
author: "Gabrielle Barsotti"
date: "11/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Summary

- In this document I will be exploring the difference between purple and red sea urchins located off the Santa Barbara Coast. The data was taken from red and purple sea urchins at 5 different locations. 

## Read in the Urchin Data

```{r message = FALSE}
urchin_data <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Exploring data visulization

  - Using the data to compare the sizes between purple and red sea urchins
  - Only need to do minimal updates to the data tables
  
```{r}
ggplot(data = urchin_data, aes(x = size)) +
  geom_histogram(bins = 15) +
  facet_wrap(~common_name) +
  labs(x = "Urchin Size",
       y = "Count")

```
  
```{r}
ggplot(data = urchin_data, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = "Common Name",
       y = "Size")

```
 
### Summary Table & Descriptive Statistics

- This summary table will look at the mean, median and standard diviation of both purple and red sea urchins

```{r message = FALSE}
urchin_stats <- urchin_data %>% 
  group_by(common_name) %>%
  summarize(mean = mean(size),
            median = median(size),
            sd = sd(size),
            sample_size = n())
            
  kbl(urchin_stats) %>% 
    kable_styling()
```

### Calculation for difference in mean
```{r}
6.776489 - 4.469797
```

### Calculation for percent difference
```{r}
(6.776489 - 4.469797) / ((4.469797 + 6.776489) / 2) *100
```

### Description

- There is a 2.31 centimeter difference in the mean diameter between the purple and red sea urchin. The mean diameter of red urchin data samples is 2.31 centimeters or a 41.02% larger than the mean diameter of the purple urchin data samples.

## Section 2 : Exploring purple urchin count by site and year

```{r message = FALSE}
urchin_clean_dates <- urchin_data %>% 
  mutate(date = mdy(date))

class(urchin_clean_dates$date)
  
purple_urchin <- urchin_clean_dates %>%
  filter(common_name == "Purple Urchin") %>%
  mutate(date = ymd(date)) %>% 
  mutate(year = year(date)) %>%
  mutate(site_full = recode(site, "NAPL" = "Naples Reef", "AQUE" = "Arroyo Quemado Reef", "CARP" = "Carpinteria Reef",
                            "MOHK" = "Mohawk Reef", "IVEE" = "Isla Vista Reef"))
  
```
 
### Summary table 
 
- This table contains the total number of purple urchins observed by site and year.

```{r}
purple_urchin_total <- purple_urchin %>%
  select("site_full", "year") %>% 
  group_by(site_full, year) %>% 
  count(site_full)
  
```

### Made an exploritory graph 

- This graph shows the counts of purple urchins recorded at each site over time for years included in the summary table 

```{r}
ggplot(data = purple_urchin_total, aes(x = year, y = n)) +
  geom_col(aes(color = site_full)) +
  facet_wrap(~site_full) +
  labs(x = "Count",
       y = "Year",
       title = "Annual Purple Sea Urchin Count by Location_Barsotti")
```




